---
title: "TBI - Single nuclei RNAseq"
author: "Raquel Garza"
output:
  html_document:
    df_print: paged
---

This Markdown encompass the code that was used to generate the gene expression related figures of the (manuscript)[https://doi.org/10.1101/2022.09.07.506982] titled: Single-cell transcriptomics of resected human traumatic brain injury tissues reveals acute activation of endogenous retroviruses in oligodendroglia.

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Volumes/MyPassport/TBI/03.05.22/')
```

```{r}
set.seed(7)
library(ggplot2)
library(Seurat)
library(ggpubr)
library(biomaRt)
library(pheatmap)
library(RColorBrewer)
library(stringr)
library(openxlsx)
library(dplyr)
library(clusterProfiler)
library(org.Hs.eg.db)
library(AnnotationHub)

tbi <- readRDS("results-merged.rds")
tbi_seurat <- tbi[[1]]
tbi_res <- tbi[[2]]

getPalette = colorRampPalette(brewer.pal(9, "Greys"))
DimPlot(tbi_seurat, cols = getPalette(23)[5:20], label = T) + theme(legend.position = "None")

```

### Just to have the sample names at hand
```{r}
tbi_samples <- c("MJ_TBI_nr1_LT", "MJ_TBI_nr10_LT", "MJ_TBI_nr11_LFP", "MJ_TBI_nr16_RT", "MJ_TBI_nr19_RF", "MJ_TBI_nr2_LF", "MJ_TBI_nr20_RF", "MJ_TBI_nr21_RF", "MJ_TBI_nr3_RF", "MJ_TBI_nr6_RT", "MJ_TBI_nr7_LT", "MJ_TBI_nr8_RT")
control_samples <- c("TBI_HuBrainCTL_Nuclei501F_F", "TBI_HuBrainCTL_Nuclei501T_T", "TBI_HuBrainCTL_Nuclei502T_T", "TBI_HuBrainCTL_Nuclei529F_F", "TBI_HuBrainCTL_Nuclei529T_T")
```

### Some relevant markers
```{r}
DotPlot(tbi_seurat,  features=c("MAP2", "DCX", "RBFOX1", "GRIN1", "HS3ST2", "GAD1", "GAD2", "CALB2", "CNR1", "GFAP", "AQP4", "GJA1", "SLC1A3", "PLP1", "MOG", "COL9A1", "VCAN", "PDGFRA", "P2RY12", "FYB1")) + coord_flip()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="")

FeaturePlot(tbi_seurat, c("RBFOX3", "GAD1", "GFAP", "PLP1", "VCAN", "FYB1"), ncol=3, cols = c("lightgrey", "red"))
```

### Add cell types to metadata
```{r}
celltypes <- c("5" = "Excitatory Neurons",
  "0" = "Excitatory Neurons",
  "6" = "Excitatory Neurons",
  "2" = "Interneurons",
  "4" = "Astrocytes",
  "9" = "Interneurons",
  "3" = "Interneurons",
  "1" = "Oligodendrocytes",
  "8" = "Microglia",
  "7" = "OPC",
  "14" = "Endothelial",
  "13" = "Endothelial",
  "10" = "?",
  "11" = "?",
  "12" = "?")

celltypes <- as.data.frame(celltypes)
celltypes$seurat_clusters <- rownames(celltypes)

clusters <- FetchData(tbi_seurat, "seurat_clusters")
clusters$barcode <- rownames(clusters)
clusters <- merge(clusters, celltypes, by="seurat_clusters")
colnames(clusters) <- c("seurat_clusters", "barcode", "celltypes")
rownames(clusters) <- clusters$barcode

tbi_seurat <- AddMetaData(tbi_seurat, metadata = clusters[,"celltypes", drop=F], col.name = "celltypes")

DimPlot(tbi_seurat, group.by = "celltypes",cols = c("Excitatory Neurons" = "#6d68a1",
                               "Interneurons" = "#312c6e",
                               "Oligodendrocytes" = "#579160",
                               "Astrocytes" = "#2b6b35",
                               "OPC" = "#104d1a",
                               "Microglia" = "#e8cc4d",
                               "Endothelial" = "#e3a94b",
                               "?" = "#c2831d"), label = T, label.box = T,  label.color = "white") + ggtitle("")
```


### Add condition to metadata
```{r}
clusters <- FetchData(tbi_seurat, "sample")
clusters$condition <- ifelse(clusters$sample %in% tbi_samples, "TBI", "Control")
tbi_seurat <- AddMetaData(tbi_seurat, metadata = clusters[,"condition", drop=F], col.name = "condition")

DimPlot(tbi_seurat,  label = T,  split.by = "condition") 
```

### Num of nuclei per sample
```{r}
sample_celltypes <- as.data.frame(table(tbi_seurat$sample, tbi_seurat$celltypes))
colnames(sample_celltypes) <- c("sample", "celltype", "Freq")
# Import metadata
tbi_metadata <- read.xlsx("samples_metadata.xlsx")
sample_celltypes <- merge(sample_celltypes, tbi_metadata, by.x="sample", by.y="ID")
sample_celltypes$key_condition <- paste(sample_celltypes$Condition, sample_celltypes$key, sep=" - ")

ggplot(sample_celltypes, aes(x=key_condition, y=Freq, fill=Condition)) + geom_bar(stat="identity", position = "dodge")+ facet_wrap(.~celltype, scales= "free", ncol=4) + theme_classic() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(x="", y="Percentage of nuclei in cell type (%)")+ scale_fill_manual(values = c("#bab8b8", "#db8f8f"))
```

### Number of nuclei per cell type per condition
```{r}
conditions_celltypes <- FetchData(tbi_seurat, c("condition", "celltypes"))
conditions_celltypes <- as.data.frame(table(conditions_celltypes$condition, conditions_celltypes$celltypes))
conditions_celltypes$Var2 <- factor(conditions_celltypes$Var2, levels = c("Excitatory Neurons", "Interneurons", "Oligodendrocytes", "Astrocytes", "OPC", "Microglia", "?", "Endothelial"))

total_num_cells_conditions <- table(FetchData(tbi_seurat, c("condition")))

conditions_celltypes$Freq <- conditions_celltypes$Freq/(ifelse(conditions_celltypes$Var1 == "Control", total_num_cells_conditions[["Control"]], total_num_cells_conditions[["TBI"]]))

ggplot(conditions_celltypes, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity") + 
  theme_classic() + labs(x="", y="Ratio of cells", fill="") + ylim(c(0,1)) + 
  scale_fill_manual(values = c("Excitatory Neurons" = "#6d68a1",
                               "Interneurons" = "#312c6e",
                               "Oligodendrocytes" = "#579160",
                               "Astrocytes" = "#2b6b35",
                               "OPC" = "#104d1a",
                               "Microglia" = "#e8cc4d",
                               "Endothelial" = "#e3a94b",
                               "?" = "#c2831d")) + theme(text = element_text(size=15))

sample_celltypes <- FetchData(tbi_seurat, c("sample", "celltypes"))
sample_celltypes <- as.data.frame(table(sample_celltypes$sample, sample_celltypes$celltypes))
sample_celltypes$Var2 <- factor(sample_celltypes$Var2, levels = c("Excitatory Neurons", "Interneurons", "Oligodendrocytes", "Astrocytes", "OPC", "Microglia", "?", "Endothelial"))

ggplot(sample_celltypes, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position = "fill") + 
  theme_classic() + labs(x="", y="Ratio of cells", fill="") +
  scale_fill_manual(values = c("Excitatory Neurons" = "#6d68a1",
                               "Interneurons" = "#312c6e",
                               "Oligodendrocytes" = "#579160",
                               "Astrocytes" = "#2b6b35",
                               "OPC" = "#104d1a",
                               "Microglia" = "#e8cc4d",
                               "Endothelial" = "#e3a94b",
                               "?" = "#c2831d")) + theme(text = element_text(size=15)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(sample_celltypes, aes(x=Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity") + 
  theme_classic() + labs(x="", y="Number of cells", fill="") +
  scale_fill_manual(values = c("Excitatory Neurons" = "#6d68a1",
                               "Interneurons" = "#312c6e",
                               "Oligodendrocytes" = "#579160",
                               "Astrocytes" = "#2b6b35",
                               "OPC" = "#104d1a",
                               "Microglia" = "#e8cc4d",
                               "Endothelial" = "#e3a94b",
                               "?" = "#c2831d")) + theme(text = element_text(size=15)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

### Cell cycle scoring
```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
tbi_seurat <- CellCycleScoring(tbi_seurat, s.features = s.genes, g2m.features = g2m.genes, set.ident = F, assay="RNA")
tbi_seurat <- AddMetaData(tbi_seurat, metadata = ifelse(tbi_seurat$Phase == "G1", "non-cycling", "cycling"), col.name = "cellCycle")

cellcycle_scores <- FetchData(tbi_seurat, c("UMAP_1", "UMAP_2","condition", "G2M.Score", "S.Score"))

ggarrange(ggplot(cellcycle_scores, aes(x=UMAP_1, y=UMAP_2, colour=G2M.Score)) + geom_point(size=0.5) +
            facet_wrap(.~condition) + theme_classic() + scale_colour_gradient(low="lightgrey", high = "red"),
          ggplot(cellcycle_scores, aes(x=UMAP_1, y=UMAP_2, colour=S.Score)) + geom_point(size=0.5) + 
            facet_wrap(.~condition) + theme_classic() + scale_colour_gradient(low="lightgrey", high = "red"), ncol=1)

```

### STAT1, STAT2, CD47, NLRC5 featureplots
```{r}
stat1_control <- FeaturePlot(subset(tbi_seurat, condition == "Control"), features = c("STAT1"))
stat1_tbi <- FeaturePlot(subset(tbi_seurat, condition == "TBI"), features = c("STAT1"))
stat1_control$data$condition <- "Healthy"
stat1_tbi$data$condition <- "TBI"
stat1 <- rbind(stat1_control$data, 
               stat1_tbi$data)
colnames(stat1)[4] <- "STAT1"
ggplot(stat1, aes(x=UMAP_1, y=UMAP_2, colour=(STAT1))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("STAT1")

stat2_control <- FeaturePlot(subset(tbi_seurat, condition == "Control"), features = c("STAT2"))
stat2_tbi <- FeaturePlot(subset(tbi_seurat, condition == "TBI"), features = c("STAT2"))
stat2_control$data$condition <- "Healthy"
stat2_tbi$data$condition <- "TBI"
stat2 <- rbind(stat2_control$data, 
               stat2_tbi$data)
colnames(stat2)[4] <- "STAT2"
ggplot(stat2, aes(x=UMAP_1, y=UMAP_2, colour=(STAT2))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("STAT2")

cd47_control <- FeaturePlot(subset(tbi_seurat, condition == "Control"), features = c("CD47"))
cd47_tbi <- FeaturePlot(subset(tbi_seurat, condition == "TBI"), features = c("CD47"))
cd47_control$data$condition <- "Healthy"
cd47_tbi$data$condition <- "TBI"
cd47 <- rbind(cd47_control$data, 
               cd47_tbi$data)
colnames(cd47)[4] <- "CD47"
ggplot(cd47, aes(x=UMAP_1, y=UMAP_2, colour=(CD47))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("CD47")

nlrc5_control <- FeaturePlot(subset(tbi_seurat, condition == "Control"), features = c("NLRC5"))
nlrc5_tbi <- FeaturePlot(subset(tbi_seurat, condition == "TBI"), features = c("NLRC5"))
nlrc5_control$data$condition <- "Healthy"
nlrc5_tbi$data$condition <- "TBI"
nlrc5 <- rbind(nlrc5_control$data, 
               nlrc5_tbi$data)
colnames(nlrc5)[4] <- "NLRC5"
ggplot(nlrc5, aes(x=UMAP_1, y=UMAP_2, colour=(NLRC5))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("NLRC5")

ciita_control <- FeaturePlot(subset(tbi_seurat, condition == "Control"), features = c("CIITA"))
ciita_tbi <- FeaturePlot(subset(tbi_seurat, condition == "TBI"), features = c("CIITA"))
ciita_control$data$condition <- "Healthy"
ciita_tbi$data$condition <- "TBI"
ciita <- rbind(ciita_control$data, 
               ciita_tbi$data)
colnames(ciita)[4] <- "CIITA"
ggplot(ciita, aes(x=UMAP_1, y=UMAP_2, colour=(CIITA))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("CIITA")

```


### Annotate time post injury

Cell type composition 
```{r}
rownames(tbi_metadata) <- tbi_metadata$ID
time <- tbi_metadata[,"Time_Post_Injury",drop=F]
time$sample <- rownames(time)

time_sample <- FetchData(tbi_seurat, "sample")
time_sample$barcode <- rownames(time_sample)
time_sample <- merge(time_sample, time, by="sample")
rownames(time_sample) <- time_sample$barcode

tbi_seurat <- AddMetaData(tbi_seurat, metadata = time_sample[,"Time_Post_Injury", drop=F], col.name = "time")
FeaturePlot(subset(tbi_seurat, condition == "TBI"), "time")
```

### Annotate region of injury
```{r}
regions <- tbi_metadata[,"region",drop=F]
regions$sample <- rownames(regions)

regions_sample <- FetchData(tbi_seurat, "sample")
regions_sample$barcode <- rownames(regions_sample)
regions_sample <- merge(regions_sample, regions, by="sample")
rownames(regions_sample) <- regions_sample$barcode

tbi_seurat <- AddMetaData(tbi_seurat, metadata = regions_sample[,"region", drop=F], col.name = "region")

DimPlot(tbi_seurat, group.by = "region") + ggtitle("Regions")
```

### GSEA - Pooling clusters of same cell types together
```{r}
set.seed(6)
tbi_seurat <- SetIdent(tbi_seurat, value="celltypes")
celltypes <- as.character(unique(Idents(tbi_seurat)))
deas_celltype <- sapply(as.character(celltypes), function(x) NULL)
deas_celltype_filtered <- sapply(as.character(celltypes), function(x) NULL)
for(celltype in as.character(celltypes)) deas_celltype[[celltype]] <- FindMarkers(tbi_seurat, group.by = "condition", ident.1 = "TBI", subset.ident = celltype)
for(celltype in as.character(celltypes)) deas_celltype[[celltype]]$gene <- rownames(deas_celltype[[celltype]])
for(celltype in as.character(celltypes)) rownames(deas_celltype[[celltype]]) <- NULL
for(celltype in as.character(celltypes)) deas_celltype_filtered[[celltype]] <- deas_celltype[[celltype]][which(deas_celltype[[celltype]]$p_val_adj < 0.01),]

gse_dotplot <- function(dea){
  genelist <- bitr(rownames(dea), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
  genelist <- merge(genelist, dea[,c("avg_log2FC"), drop=F], by.x="SYMBOL", by.y="row.names")
  genelist <- genelist[order(genelist$avg_log2FC, decreasing = T),]
  
  genelist_FC <- genelist$avg_log2FC
  names(genelist_FC) <- genelist$ENTREZID
  gse <- gseGO(geneList=genelist_FC, 
               ont ="ALL", 
               keyType = "ENTREZID", 
               minGSSize = 3, 
               maxGSSize = 800, 
               seed = T,
               pvalueCutoff = 0.05,
               verbose = TRUE, 
               OrgDb = org.Hs.eg.db, 
               pAdjustMethod = "BH")
  return(gse)
}

gses_celltype <- sapply(as.character(celltypes), function(x) NULL)
for(celltype in as.character(celltypes)) rownames(deas_celltype[[celltype]]) <- deas_celltype[[celltype]]$gene
for(celltype in as.character(celltypes)) rownames(deas_celltype_filtered[[celltype]]) <- deas_celltype_filtered[[celltype]]$gene

gses_celltype_oligo <- read.xlsx("tables/GO_overrepresentation/seed_7/per_celltype/oligodendrocytes.xlsx")
gses_celltype_oligo$sign <- ifelse(gses_celltype_oligo$enrichmentScore < 0, "Supressed", "Activated")
gses_celltype_oligo$num_genes <- sapply(str_split(gses_celltype_oligo$core_enrichment, "/"), length)
gses_celltype_oligo$gene_ratio <- gses_celltype_oligo$num_genes / gses_celltype_oligo$setSize
gses_celltype_oligo$Description <- factor(gses_celltype_oligo$Description, levels = gses_celltype_oligo[order(gses_celltype_oligo$gene_ratio),"Description"])
gses_celltype_oligo <- gses_celltype_oligo[which(gses_celltype_oligo$ONTOLOGY == "BP"),]
gses_celltype_oligo$celltype <- "oligodendrocytes"

gses_celltype_opc <- read.xlsx("tables/GO_overrepresentation/seed_7/per_celltype/OPC.xlsx")
gses_celltype_opc$sign <- ifelse(gses_celltype_opc$enrichmentScore < 0, "Supressed", "Activated")
gses_celltype_opc$num_genes <- sapply(str_split(gses_celltype_opc$core_enrichment, "/"), length)
gses_celltype_opc$gene_ratio <- gses_celltype_opc$num_genes / gses_celltype_opc$setSize
gses_celltype_opc$Description <- factor(gses_celltype_opc$Description, levels = gses_celltype_opc[order(gses_celltype_opc$gene_ratio),"Description"])
gses_celltype_opc <- gses_celltype_opc[which(gses_celltype_opc$ONTOLOGY == "BP"),]
gses_celltype_opc$celltype <- "OPC"

gses_celltype <- rbind(gses_celltype,
                       gses_celltype_opc)

insert_new_lines <- function(str, n) {gsub(paste0("([^ ]+( +[^ ]+){",n-1,"}) +"),
                              "\\1\n", str)}
gses_celltype$Description <- insert_new_lines(gses_celltype$Description, 3)
gses_celltype$Description <- factor(gses_celltype$Description, levels = unique(gses_celltype[order(gses_celltype$celltype, gses_celltype$gene_ratio),"Description"]))
# pdf("gsea_oligodendroglia.pdf", height=7, width = 12)
ggplot(gses_celltype, aes(x=Description, y=gene_ratio, size = num_genes, fill = p.adjust)) + geom_point(shape=21, colour="lightgrey") + facet_wrap(.~sign+celltype) + coord_flip() + theme_pubr(legend = "right", border = T) + labs(x="", fill="Padj", y = "Gene ratio", size = "Num genes") + scale_fill_gradientn(colors = c("firebrick1", "gray94")) + scale_size_continuous(range = c(2,15)) + theme(axis.text.y = element_text(size=15),
                                                                                                                                                                                                                                                                                                                                                                                                            strip.text.x = element_text(size = 15),
                                                                                                                                                                                                                                                                                                                                                                                                            axis.text.x = element_text(size=15),
                                                                                                                                                                                                                                                                                                                                                                                                            legend.text = element_text(size=15),
                                                                                                                                                                                                                                                                                                                                                                                                            legend.title = element_text(size=16)) 
# dev.off()

# Saved the results to xlsx at tables/GO_overrepresentation/seed_7/per_celltype/
# gses_celltype$`Excitatory Neurons` <- gse_dotplot(deas_celltype_filtered$`Excitatory Neurons`)
# gses_celltype$Interneurons <- gse_dotplot(deas_celltype_filtered$Interneurons)
# gses_celltype$Oligodendrocytes <- gse_dotplot(deas_celltype_filtered$Oligodendrocytes)
# gses_celltype$OPC <- gse_dotplot(dea=deas_celltype_filtered$OPC)
# gses_celltype$Microglia <- gse_dotplot(dea=deas_celltype_filtered$Microglia)
# gses_celltype$Astrocytes <- gse_dotplot(deas_celltype_filtered$Astrocytes)
# gses_celltype$Endothelial <- gse_dotplot(deas_celltype_filtered$Endothelial) 
# gses_celltype$`?` <- gse_dotplot(deas_celltype_filtered$`?`)
```
Saved the results to xlsx at tables/GO_overrepresentation/seed_7/per_celltype/

### Overrepresented terms in oligodendrocytes
```{r}
gses_celltype_oligos <- read.xlsx("tables/GO_overrepresentation/seed_7/per_celltype/oligodendrocytes.xlsx")

response_genes_oligos <- gses_celltype_oligos[which(gses_celltype_oligos$enrichmentScore > 0),]
# response_genes_oligos <- gses_celltype_oligos[which(gses_celltype_oligos$Description == "interferon-gamma-mediated signaling pathway"),]
response_genes_oligos <- unique(unlist(str_split(response_genes_oligos$core_enrichment, "/")))
response_genes_oligos <- bitr(response_genes_oligos, fromType="ENTREZID", toType="SYMBOL", OrgDb="org.Hs.eg.db", drop = F)

response_genes_oligodendroglia <- unique(c(response_genes_oligos$SYMBOL))
tbi_seurat <- ScaleData(tbi_seurat, rownames(tbi_seurat))
gene.data <- response_genes_oligodendroglia[which(response_genes_oligodendroglia %in% c(deas_celltype_filtered$Oligodendrocytes[which(deas_celltype_filtered$Oligodendrocytes$avg_log2FC > 0),"gene"]))]

tmp <- DoHeatmap(subset(tbi_seurat, celltypes %in% c("Oligodendrocytes", "OPC", "Astrocytes", "Microglia", "Excitatory Neurons", "Interneurons", "Endothelial", "?")), features = gene.data)# $hgnc_symbol
tmp_df <- reshape2::dcast(tmp$data[,1:3], formula= Feature~Cell, value.var = "Expression")
rownames(tmp_df) <- tmp_df$Feature
tmp_df <- tmp_df[,-1]
annotation_col <- FetchData(tbi_seurat, vars = c("condition", "celltypes", "time"))
annotation_col <- annotation_col[order(annotation_col$celltypes, annotation_col$condition),,drop=F]
annotation_col <- annotation_col[which(rownames(annotation_col) %in% colnames(tmp_df)),, drop=F]
annotation_col$celltypes <- factor(annotation_col$celltypes, levels = c("Oligodendrocytes", "OPC", "Astrocytes", "Microglia", "Excitatory Neurons", "Interneurons", "Endothelial", "?"))
annotation_col <- annotation_col[order(annotation_col$celltypes, annotation_col$condition, annotation_col$time),]
annotation_col$time <- as.factor(annotation_col$time)
col.pal <- colorRampPalette(colors = c("lightgrey", "white", "#2d518a"))(60)
# Takes forever to plot... 
# png("heatmap_immune.png", res = 350, width = 5000, height = 800)
# pdf("heatmap_immune.pdf", width=50)
pheatmap(tmp_df[,rownames(annotation_col)], cluster_cols = F, show_rownames = F, cluster_rows = T,  show_colnames = F, color = col.pal, annotation_col = annotation_col, 
         gaps_col = c(which(!duplicated(paste(annotation_col$condition, annotation_col$celltypes)))[-1]-1, rep(which(!duplicated(annotation_col$celltypes))[-1]-1, 3)),
         fontsize = 7, 
           annotation_colors = list("condition" = c("Control" = "#BFBFBF",
                                                    "TBI" = "#D493C0"),
                                    "celltypes" = c("Oligodendrocytes" = "#BDA6CD",
                                                    "OPC" = "#7F6791",
                                                    "Astrocytes" = "#92C876",
                                                    "Microglia" = "#ED7B29",
                                                    "Excitatory Neurons" = "#6FC7D4",
                                                    "Interneurons" = "#5DA19A",
                                                    "Endothelial" = "#D2724B",
                                                    "?" = "#96979A"),
                                    "time" = c("4" = "#BEA7CC",
                                               "9" = "#AA91BC",
                                               "24" = "#947AA9",
                                               "42" = "#72578B",
                                               "57" = "#564174",
                                               "84" = "#40305B",
                                               "106" = "#2F2247",
                                               "180" = "#1D1532",
                                               "192" = "#181425")), treeheight_row = 0)
# dev.off()

```

### Immune related enrichment

#### Enrichment score per cell type
```{r}
features <- list("immune_response_oligos" = gene.data)
tbi_seurat <- AddModuleScore(tbi_seurat, features = features, name = c("immune_response_oligos"), assay = "RNA", ctrl=5)
tmp <- VlnPlot(tbi_seurat, features = "immune_response_oligos1", group.by = "celltypes", split.by = "condition", pt.size = 0) 

# pdf("violin_enrichment_immune.pdf", width = 17, height = 3)
ggplot(tmp$data, aes(x=split, y=immune_response_oligos1, fill=split)) + geom_violin() + facet_wrap(.~ident, scales = "free_x", ncol = 8) + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
  ggmin::theme_powerpoint() + geom_boxplot(width=0.1, color="black", alpha=0.2) +
  geom_hline(yintercept = 0, linetype="dashed", colour="red") + labs(y="Enrichment score", x="")
# dev.off()
```

#### Enrichment score in oligodendrocytes per time post injury
```{r}
VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = "immune_response_oligos1", group.by = "time", pt.size = 0) + ggtitle("Immune-related genes enrichment score in oligodendrocytes") + labs(x="Time after injury") + geom_vline(xintercept = 0, linetype = "dotted", colour="red") + stat_summary(geom = "point",fun.y = "mean",col = "black",size = 3,shape = 16)
```

### Violin plots: Gene expression in oligodendrocytes (STAT1, 2, NLRC5...)
```{r}
padj_genes <- rbind(deas_celltype$Oligodendrocytes["STAT1", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["STAT2", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["IRF1", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["PARP14", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["PARP12", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["PARP9", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["NLRC5", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["CIITA", "p_val_adj", drop=F],
                    deas_celltype$Oligodendrocytes["IFI16", "p_val_adj", drop=F])
padj_genes$p_val_adj <- ifelse(padj_genes$p_val_adj > 0.05 | is.na(padj_genes$p_val_adj), "n.s.", format(padj_genes$p_val_adj, digits = 2))
library(ggpubr)
ggarrange(VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("STAT1"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red") + geom_text(x=1.5, y=3, label=padj_genes["STAT1","p_val_adj"]),
          VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("STAT2"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red") + geom_text(x=1.5, y=3, label=padj_genes["STAT2","p_val_adj"]),
          VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("NLRC5"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=3, label=padj_genes["NLRC5","p_val_adj"]), ncol=5, nrow=2)


# pdf("violin_oligodendrocytes_ifn.pdf", height = 3, width = 12)
ggarrange(VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("IRF1"), pt.size = 0.3, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=3.5, label=padj_genes["IRF1","p_val_adj"]),
          VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("IFI16"), pt.size = 0.3, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")+ geom_text(x=1.5, y=3, label=padj_genes["IFI16","p_val_adj"]),
          VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("PARP9"), pt.size = 0.3, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=3, label=padj_genes["PARP9","p_val_adj"]),
          VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("PARP12"), pt.size = 0.3, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2.3, label=padj_genes["PARP12","p_val_adj"]),
          VlnPlot(subset(tbi_seurat, celltypes == "Oligodendrocytes"), features = c("PARP14"), pt.size = 0.3, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=4, label=padj_genes["PARP14","p_val_adj"]), ncol=5, nrow=1)
# dev.off()
```

### MHC molecules in oligodendrocytes
```{r}
tbi_seurat <- ScaleData(tbi_seurat, features = rownames(tbi_seurat))
tmp <- DoHeatmap(subset(tbi_seurat, celltypes %in% c("Oligodendrocytes", "OPC")), features = c("HLA-A", "HLA-B", "HLA-C", "HLA-E", "HLA-F", "HLA-DMA", "HLA-DOB", "HLA-DRA", "HLA-DQB1", "HLA-DPA1",
                                                                                               "MYO1E",  "PSMB9", "TAPBP", "NLRC5"))
tmp_df <- reshape2::dcast(tmp$data, formula= Feature~Cell, value.var = "Expression")
rownames(tmp_df) <- tmp_df$Feature
tmp_df <- tmp_df[,-1]
annotation_col <- FetchData(tbi_seurat, vars = c("condition", "celltypes", "time"))
annotation_col <- annotation_col[order(annotation_col$condition, annotation_col$celltypes, annotation_col$time),,drop=F]
annotation_col <- annotation_col[which(rownames(annotation_col) %in% colnames(tmp_df)),, drop=F]
annotation_col$time <- as.factor(annotation_col$time)
col.pal <- RColorBrewer::brewer.pal(9, "Blues")

annotation_row <- data.frame(gene = c("HLA-A", "HLA-B", "HLA-C", "HLA-E", "HLA-F", "HLA-DMA", "HLA-DOB", "HLA-DRA", "HLA-DQB1", "HLA-DPA1", "MYO1E",  "PSMB9", "TAPBP", "NLRC5"),
           type = c(rep("molecule", 10), rep("regulator", 4)))
rownames(annotation_row) <- annotation_row$gene
annotation_row <- annotation_row[,"type",drop=F]
annotation_row <- annotation_row[order(annotation_row$type),,drop=F]
# pdf("oligodendroglia_heatmap_adaptive_immune_short.pdf")
# png("oligodendroglia_heatmap_adaptive_immune_short.png", res = 350, width = 1500, height = 800)
pheatmap(tmp_df[rownames(annotation_row),rownames(annotation_col)], cluster_cols = F, show_rownames = T, cluster_rows = F,  show_colnames = F, color = col.pal, annotation_col = annotation_col, gaps_col =which(!duplicated(annotation_col))[3]-1, fontsize = 7, gaps_row = 10,
         annotation_colors = list("condition" = c("Control" = "#BFBFBF",
                                                  "TBI" = "#D493C0"),
                                  "celltypes" = c("Oligodendrocytes" = "#BDA6CD",
                                                  "OPC" = "#7F6791"),
                                  "time" = c("4" = "#BEA7CC",
                                             "9" = "#AA91BC",
                                             "24" = "#947AA9",
                                             "42" = "#72578B",
                                             "57" = "#564174",
                                             "84" = "#40305B",
                                             "106" = "#2F2247",
                                             "180" = "#1D1532",
                                             "192" = "#181425")), treeheight_row = 0)
# dev.off()
```

### MHC moleculesin all cell types
```{r}
tmp <- DoHeatmap(tbi_seurat, features = c("HLA-A", "HLA-B", "HLA-C", "HLA-E", "HLA-F", "HLA-DMA", "HLA-DOB", "HLA-DRA", "HLA-DQB1", "HLA-DPA1"))
tmp_df <- reshape2::dcast(tmp$data, formula= Feature~Cell, value.var = "Expression")
rownames(tmp_df) <- tmp_df$Feature
tmp_df <- tmp_df[,-1]
annotation_col <- FetchData(tbi_seurat, vars = c("condition", "celltypes", "time"))
annotation_col$celltypes <- factor(annotation_col$celltypes, levels = c("Oligodendrocytes", "OPC", "Astrocytes", "Microglia", "Excitatory Neurons", "Interneurons", "Endothelial", "?"))
annotation_col <- annotation_col[order(annotation_col$celltypes,annotation_col$condition, annotation_col$time),,drop=F]
annotation_col <- annotation_col[which(rownames(annotation_col) %in% colnames(tmp_df)),, drop=F]
annotation_col$time <- as.factor(annotation_col$time)

col.pal <- RColorBrewer::brewer.pal(9, "Blues")

# png("heatmap_adaptive_immune.png", res = 350, width = 5000, height = 800)
# pdf("heatmap_adaptive_immune.pdf", width=50)
pheatmap(tmp_df[,rownames(annotation_col)], cluster_cols = F, show_rownames = F, cluster_rows = T,  show_colnames = F, color = col.pal, annotation_col = annotation_col, 
         gaps_col = c(which(!duplicated(paste(annotation_col$condition, annotation_col$celltypes)))[-1]-1, rep(which(!duplicated(annotation_col$celltypes))[-1]-1, 3)),
         fontsize = 7, 
           annotation_colors = list("condition" = c("Control" = "#BFBFBF",
                                                    "TBI" = "#D493C0"),
                                    "celltypes" = c("Oligodendrocytes" = "#BDA6CD",
                                                    "OPC" = "#7F6791",
                                                    "Astrocytes" = "#92C876",
                                                    "Microglia" = "#ED7B29",
                                                    "Excitatory Neurons" = "#6FC7D4",
                                                    "Interneurons" = "#5DA19A",
                                                    "Endothelial" = "#D2724B",
                                                    "?" = "#96979A"),
                                    "time" = c("4" = "#BEA7CC",
                                               "9" = "#AA91BC",
                                               "24" = "#947AA9",
                                               "42" = "#72578B",
                                               "57" = "#564174",
                                               "84" = "#40305B",
                                               "106" = "#2F2247",
                                               "180" = "#1D1532",
                                               "192" = "#181425")), treeheight_row = 0)
# dev.off()
```

#### Enrichment score per cell type adaptive immunity
```{r}
features <- list("hla" = c("HLA-A", "HLA-B", "HLA-C", "HLA-E", "HLA-F", "HLA-DMA", "HLA-DOB", "HLA-DRA", "HLA-DQB1", "HLA-DPA1"))
tbi_seurat <- AddModuleScore(tbi_seurat, features = features, name = c("hla"), assay = "RNA", ctrl=5)
tmp <- VlnPlot(tbi_seurat, features = "hla1", group.by = "celltypes", split.by = "condition", pt.size = 0) 

# pdf("violin_enrichment_hla.pdf", width = 17, height = 3)
ggplot(tmp$data, aes(x=split, y=hla1, fill=split)) + geom_violin() + facet_wrap(.~ident, scales = "free_x", ncol = 8) + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
  ggmin::theme_powerpoint() + geom_boxplot(width=0.1, color="black", alpha=0.2) +
  geom_hline(yintercept = 0, linetype="dashed", colour="red") + labs(y="Enrichment score", x="")
# dev.off()
```

### Violin plots: Gene expression in OPCs (STAT1, 2, NLRC5...)
```{r}
padj_genes <- rbind(deas_celltype$OPC["STAT1", "p_val_adj", drop=F],
                    deas_celltype$OPC["STAT2", "p_val_adj", drop=F],
                    deas_celltype$OPC["IRF1", "p_val_adj", drop=F],
                    deas_celltype$OPC["PARP14", "p_val_adj", drop=F],
                    deas_celltype$OPC["NLRC5", "p_val_adj", drop=F],
                    deas_celltype$OPC["CIITA", "p_val_adj", drop=F],
                    deas_celltype$OPC["IFI16", "p_val_adj", drop=F],
                    deas_celltype$OPC["PARP9", "p_val_adj", drop=F],
                    deas_celltype$OPC["PAR12", "p_val_adj", drop=F])

ggarrange(VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("STAT1"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red") + geom_text(x=1.5, y=2, label=formatC(padj_genes["STAT1","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("STAT2"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red") + geom_text(x=1.5, y=2, label=formatC(padj_genes["STAT2","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("IRF1"), pt.size = 0.1, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2, label=formatC(padj_genes["IRF1","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("PARP14"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2, label=formatC(padj_genes["PARP14","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("PARP14"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2, label=formatC(padj_genes["PARP12","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("PARP14"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2, label=formatC(padj_genes["PARP9","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("NLRC5"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2, label=formatC(padj_genes["NLRC5","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("CIITA"), pt.size = 0.1, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(x=1.5, y=2, label=formatC(padj_genes["CIITA","p_val_adj"], format = "e", digits = 2)),
          VlnPlot(subset(tbi_seurat, celltypes == "OPC"), features = c("IFI16"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")+ geom_text(x=1.5, y=2, label=formatC(padj_genes["IFI16","p_val_adj"], format = "e", digits = 2)), ncol=4, nrow=2)
```

<!-- ## From here on is the analysis without Frontal 3 - reviewers comments ### -->
<!-- ### GSEA - Pooling cell types together -->
<!-- ```{r collapse=TRUE} -->
<!-- set.seed(7) -->
<!-- tbi_wo_MJ_TBI_nr2 <- subset(tbi_seurat, sample != "MJ_TBI_nr2_LF") -->

<!-- tbi_wo_MJ_TBI_nr2 <- SetIdent(tbi_wo_MJ_TBI_nr2, value="celltypes") -->
<!-- celltypes <- as.character(unique(Idents(tbi_wo_MJ_TBI_nr2))) -->
<!-- deas_celltype <- sapply(as.character(celltypes), function(x) NULL) -->
<!-- for(celltype in as.character(celltypes)) deas_celltype[[celltype]] <- FindMarkers(tbi_wo_MJ_TBI_nr2, group.by = "conditions", ident.1 = "TBI", subset.ident = celltype) -->
<!-- for(celltype in as.character(celltypes)) deas_celltype[[celltype]]$gene <- rownames(deas_celltype[[celltype]]) -->
<!-- for(celltype in as.character(celltypes)) rownames(deas_celltype[[celltype]]) <- NULL -->
<!-- for(celltype in as.character(celltypes)) deas_celltype[[celltype]] <- deas_celltype[[celltype]][which(deas_celltype[[celltype]]$p_val_adj < 0.01),] -->

<!-- gses_celltype <- sapply(as.character(celltypes), function(x) NULL) -->
<!-- for(celltype in as.character(celltypes)) rownames(deas_celltype[[celltype]]) <- deas_celltype[[celltype]]$gene -->
<!-- gses_celltype$Endothelial <- gse_dotplot(deas_celltype$Endothelial) -->
<!-- gses_celltype$Interneurons <- gse_dotplot(deas_celltype$Interneurons) -->
<!-- gses_celltype$`Excitatory Neurons` <- gse_dotplot(deas_celltype$`Excitatory Neurons`) -->
<!-- gses_celltype$Astrocytes <- gse_dotplot(deas_celltype$Astrocytes) -->
<!-- gses_celltype$`?` <- gse_dotplot(deas_celltype$`?`) -->
<!-- gses_celltype$Oligodendrocytes <- gse_dotplot(deas_celltype$Oligodendrocytes) -->
<!-- gses_celltype$OPC <- gse_dotplot(dea=deas_celltype$OPC) # attempt to select less than one element in OneIndex -->
<!-- gses_celltype$Microglia <- gse_dotplot(dea=deas_celltype$Microglia) -->

<!-- dotplot(gses_celltype$OPC, showCategory=10, split=".sign") + facet_grid(.~.sign) + ggtitle("OPC") -->
<!-- dotplot(gses_celltype$Oligodendrocytes, showCategory=10, split=".sign") + facet_grid(.~.sign) + ggtitle("Oligodendrocytes") -->
<!-- # Save to xlsx -->

<!-- ``` -->

<!-- ### Immune response genes wo Frontal 3 -->
<!-- ```{r collapse=TRUE } -->
<!-- tbi_wo_MJ_TBI_nr2 <- ScaleData(tbi_wo_MJ_TBI_nr2, response_genes_oligos$SYMBOL) -->

<!-- tmp <- DoHeatmap(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = response_genes_oligos$SYMBOL) -->
<!-- tmp_df <- reshape2::dcast(tmp$data[,1:3], formula= Feature~Cell, value.var = "Expression") -->
<!-- rownames(tmp_df) <- tmp_df$Feature -->
<!-- tmp_df <- tmp_df[,-1] -->
<!-- annotation_col <- FetchData(tbi_wo_MJ_TBI_nr2, vars = c("condition", "time", "sample")) -->
<!-- annotation_col <- annotation_col[order(annotation_col$condition, annotation_col$time),,drop=F] -->
<!-- annotation_col <- annotation_col[which(rownames(annotation_col) %in% colnames(tmp_df)),, drop=F] -->
<!-- annotation_col$time <- as.factor(annotation_col$time) -->

<!-- col.pal <- colorRampPalette(colors = c("lightgrey", "white", "#2d518a"))(50)#RColorBrewer::brewer.pal(9, "Reds") -->
<!-- pheatmap(tmp_df[row_order, rownames(annotation_col)], cluster_cols = F, cluster_rows = F,  show_colnames = F, color = col.pal, annotation_col = annotation_col[,c("time"), drop=F]) -->
<!-- ``` -->

<!-- ### Gene expression in oligodendrocytes without sample Frontal 3 -->
<!-- ```{r collapse=TRUE} -->
<!-- dea_oligos_complete <- FindMarkers(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), group.by = "condition", ident.1 = "TBI", lfcThreshold = 0, min.cells.feature = 1, min.cells.group = 1, min.pct = 0, random.seed = 7) -->
<!-- dea_oligos_complete$gene_name <- rownames(dea_oligos_complete) -->
<!-- padj_genes <- rbind(dea_oligos_complete["STAT1", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["STAT2", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["CD47", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["IRF1", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["PARP14", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["NLRC5", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["CIITA", "p_val_adj", drop=F], -->
<!--                     dea_oligos_complete["IFI6", "p_val_adj", drop=F]) -->

<!-- ggarrange(VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("STAT1"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red") + geom_text(aes(x="Control", y=2, label=formatC(padj_genes["STAT1","p_val_adj"], format = "e", digits = 2))), -->
<!--           VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("STAT2"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red") + geom_text(aes(x="Control", y=2, label=formatC(padj_genes["STAT2","p_val_adj"], format = "e", digits = 2))), -->
<!--           VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("IRF1"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(aes(x="Control", y=2, label=formatC(padj_genes["IRF1","p_val_adj"], format = "e", digits = 2))), -->
<!--           VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("PARP14"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(aes(x="Control", y=2, label=formatC(padj_genes["PARP14","p_val_adj"], format = "e", digits = 2))), -->
<!--           VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("NLRC5"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(aes(x="Control", y=2, label=formatC(padj_genes["NLRC5","p_val_adj"], format = "e", digits = 2))), -->
<!--           VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("CIITA"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")  + geom_text(aes(x="Control", y=2, label=formatC(padj_genes["CIITA","p_val_adj"], format = "e", digits = 2))), -->
<!--           VlnPlot(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("IFI6"), pt.size = 0, group.by = "condition") + theme(legend.position = "None") + stat_summary(fun = mean, geom='point', size = 1, color="red")+ geom_text(aes(x="Control", y=2, label=formatC(padj_genes["IFI6","p_val_adj"], format = "e", digits = 2))), ncol=4, nrow=2) -->
<!-- ``` -->

<!-- ### MHC molecules without sample Frontal 3 -->
<!-- ```{r collapse=TRUE} -->
<!-- tmp <- DoHeatmap(subset(tbi_wo_MJ_TBI_nr2, celltypes == "OPC"), features = c("HLA-A", "HLA-B", "HLA-C", "HLA-E", "HLA-F", "HLA-DMA", "HLA-DOB", "HLA-DRA", "HLA-DQB1", "HLA-DPA1")) -->
<!-- tmp_df <- reshape2::dcast(tmp$data[,1:3], formula= Feature~Cell, value.var = "Expression") -->
<!-- rownames(tmp_df) <- tmp_df$Feature -->
<!-- tmp_df <- tmp_df[,-1] -->
<!-- annotation_col <- FetchData(tbi_wo_MJ_TBI_nr2, vars = c("condition", "time")) -->
<!-- annotation_col <- annotation_col[order(annotation_col$condition, annotation_col$time),,drop=F] -->
<!-- annotation_col <- annotation_col[which(rownames(annotation_col) %in% colnames(tmp_df)),, drop=F] -->
<!-- annotation_col$time <- as.factor(annotation_col$time) -->
<!-- col.pal <- RColorBrewer::brewer.pal(9, "Reds") -->

<!-- pheatmap(tmp_df[c("HLA-A", "HLA-B", "HLA-C", "HLA-E", "HLA-F", "HLA-DMA", "HLA-DOB", "HLA-DRA", "HLA-DQB1", "HLA-DPA1"), rownames(annotation_col)], cluster_cols = F, cluster_rows = F,  show_colnames = F, color = col.pal, annotation_col = annotation_col) -->
<!-- ``` -->

<!-- ### MHC regulators  without sample Frontal 3 -->
<!-- ```{r collapse=TRUE} -->
<!-- tmp <- DoHeatmap(subset(tbi_wo_MJ_TBI_nr2, celltypes == "Oligodendrocytes"), features = c("MYO1E", "KIF5B", "PIP5K1A", "PSMB8", "PSMB9", "TAP1", "TAP2", "CD86", "CD80", "CD74", "CIITA", "NLRC5")) -->
<!-- tmp_df <- reshape2::dcast(tmp$data[,1:3], formula= Feature~Cell, value.var = "Expression") -->
<!-- rownames(tmp_df) <- tmp_df$Feature -->
<!-- tmp_df <- tmp_df[,-1] -->
<!-- annotation_col <- FetchData(tbi_wo_MJ_TBI_nr2, vars = c("condition", "time")) -->
<!-- annotation_col <- annotation_col[order(annotation_col$condition, annotation_col$time),,drop=F] -->
<!-- annotation_col <- annotation_col[which(rownames(annotation_col) %in% colnames(tmp_df)),, drop=F] -->
<!-- annotation_col$time <- as.factor(annotation_col$time) -->

<!-- pheatmap(tmp_df[c("MYO1E", "KIF5B", "PIP5K1A", "PSMB8", "PSMB9", "TAP1", "TAP2", "CD86", "CD80", "CD74", "CIITA", "NLRC5"), rownames(annotation_col)], cluster_cols = F, cluster_rows = F,  show_colnames = F, color = col.pal, annotation_col = annotation_col) -->
<!-- ``` -->

<!-- ### Gene expression without sample Frontal 3 -->
<!-- ```{r collapse=TRUE} -->
<!-- stat1_control <- FeaturePlot(subset(tbi_wo_MJ_TBI_nr2, condition == "Control"), features = c("STAT1")) -->
<!-- stat1_tbi <- FeaturePlot(subset(tbi_wo_MJ_TBI_nr2, condition == "TBI"), features = c("STAT1")) -->
<!-- stat1_control$data$condition <- "Healthy" -->
<!-- stat1_tbi$data$condition <- "TBI" -->
<!-- stat1 <- rbind(stat1_control$data,  -->
<!--                stat1_tbi$data) -->
<!-- colnames(stat1)[4] <- "STAT1" -->
<!-- ggplot(stat1, aes(x=UMAP_1, y=UMAP_2, colour=(STAT1))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("STAT1") -->

<!-- stat2_control <- FeaturePlot(subset(tbi_wo_MJ_TBI_nr2, condition == "Control"), features = c("STAT2")) -->
<!-- stat2_tbi <- FeaturePlot(subset(tbi_wo_MJ_TBI_nr2, condition == "TBI"), features = c("STAT2")) -->
<!-- stat2_control$data$condition <- "Healthy" -->
<!-- stat2_tbi$data$condition <- "TBI" -->
<!-- stat2 <- rbind(stat2_control$data,  -->
<!--                stat2_tbi$data) -->
<!-- colnames(stat2)[4] <- "STAT2" -->
<!-- ggplot(stat2, aes(x=UMAP_1, y=UMAP_2, colour=(STAT2))) + geom_point(size=0.5, alpha=0.5) + facet_wrap(.~condition) + theme_classic() + scale_color_gradient(low = "lightgray", high = "red") + labs(colour="") + theme(text=element_text(size=15)) + ggtitle("STAT2") -->
<!-- ``` -->
